---
title: "L11: Tibble Joins"
author: "Jerry"
date: "2022-10-18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tibble Joins

The idea that, given two or more tables, how would we go about putting the info into one table!

```{r}
library(tidyverse)
```

Below are two tibbles.

```{r}
# First tibble
df1 <- tibble(ID = 1:3,                     
              Name = c("Sophie", "Josh","Alex"))

# Second tibble
df2 <- tibble(ID = 2:4,                      
              Age = c(20,50,31))
```

Tibble joins require that both dataframe have a column with unique identifiers (e.g student number, address, etc.) It will then use that ID column and match values from the left or right tibble to the other.

## Mutating Tibble Joins

Mutating joins will create new tibbles.

### left_join()

`left_join()` will take all columns from the left tibble and then use the ID column to **match** values from the right tibble.

```{r}
left_join(df1, df2, by = "ID")
```

### right_join()

`right_join()` does the same but matches values on the left tibble to the right tibble.

```{r}
right_join(df1, df2, by = "ID")
```

### inner_join()

`inner_join()` retains rows only present in both sets.

```{r}
inner_join(df1, df2, by = "ID")
```

### full_join()

`full_join` retains all values and all rows.

```{r}
full_join(df1, df2, by = "ID")
```

## Filtering Joins

### semi_join()
`semi_join()` retains all rows in df1 that have a match in df2, the right tibble. These functions do not add dfs together.
```{r}
semi_join(df1, df2, by = "ID")
```


### anti_join()
`anti_join()` retains all rows in df1 that **do not** have a match in df2. 

```{r}
anti_join(df1, df2, by = "ID")
```

## Binding

### bind_rows()
`bind_rows()` appends df2 to df1 as new rows (row-wise).
```{r}
bind_rows(df1, df2)
```


### bind_cols()
`bind_cols()` appends df2 to df1 as new cols (col-wise).
```{r}
bind_cols(df1, df2)
```


## Joining Multiple Tibbles (>2)
  
Create a third tibble
```{r}
df3 <- tibble(ID = 1:5,                      
              Height = c(175,167,190,155,160))
```

You can pipe the commands together:
```{r}
full_join(df1, df2, by = "ID") %>% 
  full_join(df3, by = "ID")
```


## Joining Tibbles on Multiple Conditions
Create two new tibbles. 
```{r}
df4 <- tibble(FirstName = c("Sophie", "Josh","Alex"),
              LastName=c("Wang","Smith","Smith"),
              Age = c(42,20,50))

df5 <- tibble(First_name = c("Josh","Alex","Sophie"),        
              Last_name=c("Smith","Smith","Jones"),
              Height = c(167,190,155))

```
Above, FirstName is in df4 and First_name is in df5. They are different. 

Here's how we can specify the mapping in `full_join()`:
```{r}
full_join(df4, df5, by = c("FirstName" = "First_name", "LastName" = "Last_name"))
```

### What if we did not realize that multiple people share a last name. 
What would happen is tha function will spit out every possible combination of first name and last name (?).
Often, if you are not using a truly unique identifier in the ID paranater, you will get many more rows than columns. 


## Set Operations
Create some new example tibbles
```{r}
# First tibble
df6 <- tibble(Number = 1:3,                     
              Letter = c("A", "B","C"))

# Second tibble
df7 <- tibble(Number = 2:4,                      
              Letter = c("B","C","D"))
```


### intersect()
`intersect()` includes rows that appear in both tibbles
```{r}
intersect(df6, df7)
```

### union()
`unite()` includes rows that appear in either or both tibbles. 
```{r}
union(df6, df7)
```

### setdiff()
`setdiff()` includes rows that appear in one df but not the other. 
```{r}
setdiff(df6, df7)
```

