---
title: "L9 - Model Fitting in R"
author: "Jerry"
date: "2022-10-06"
output: html_document
---

# Model Fitting in R

Today we are going to learn basic model fitting with the relatively new package "tidymodels". If you haven't already, install the "todymodels" package.

## Load your packages

```{r}
library(tidyverse)
library(tidymodels)
library(datateachr)
library(gapminder)
```

## Load your data

```{r}
data <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", sep = ";")

data
```

## Need for Modelling Data

1.  Understand the relationship between two variables in your data.
2.  Predict the outcome of a variable given information about other variables

## Fitting Models in R

**General syntax is:**

    method(formula, data, options)

**Formulas usually look like this:**

    y ~ x1 + x2

where the tilde is more or less an equals sign.

## Let's Start Modelling

```{r}
# A simple linear regression model modelling the relationship between pH and acidity
my_lm <- lm(pH ~ fixed.acidity, data = data)
my_lm
```

## Functions Designed for Modelling in Base R

```{r}
# Grab the coefficients of the model using coef()
coef(my_lm)

# Obtain the standard error of model using sigma()
sigma(my_lm)

# Get a summary for your model using summary()
sum_lm <- summary(my_lm)

# To extract specific information out of the summary, we can using the dollar sign subset symbol
sum_lm$residuals
sum_lm$xlevels

# Get the column names of all elements of the model (which is now a list)? Basically, extract information directly from the model.
names(my_lm)
```

## Model Proving Using tidymodel: broom package

There are 3 primary functions: 1. tidy 2. augment\
3. glance

### tidy()

```{r}
# The tidy() function is analagous to the summary function but returns an output as a tibble
tidy(my_lm)
```

### augment()

```{r}
# The augment() function does something... who know what tho!
augment(my_lm) %>% 
  print(n = 5)
```

#### Predicting On New Data

```{r}
augment(my_lm, newdata = tibble(fixed.acidity = 8:13))
# for the output, it is basically saying that if fixed.acidity is x, then the pH will be y
```

### glance()

```{r}
# glance() extracts something???
glance(my_lm)
```

## predict() when there is no `broom` compatability

Not all model methods have `broom` compatabilty. `predict()` is used for these methods to probe and apply the model on the new data.

```{r}
predict(my_lm) %>% 
  unname() %>% # remove all the names of elements
  head(5)
  
```

## Plotting Models with geom_smooth()

We can plot models using geom_smooth()!

```{r}
# with no trendline
data %>% 
  ggplot(aes(x = fixed.acidity, y = pH)) +
  geom_point()

# default model
data %>% 
  ggplot(aes(x = fixed.acidity, y = pH)) +
  geom_point() +
  geom_smooth()

#linear regression model
data %>% 
  ggplot(aes(x = fixed.acidity, y = pH)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Modelling in `gapminder`

```{r}
# How does lifeExp chane over year ins
gapminder_Zimbabwe <- gapminder %>% 
  filter(country == "Zimbabwe")

# This is what our data looks like
gapminder_Zimbabwe %>% 
  ggplot(aes(year, lifeExp)) +
  geom_point()

# What model shoulw we use, a linear one probably isn't appropriate
# Let's use a quadratice model. Here, we have to specify a formula for our lm method
gapminder_Zimbabwe %>% 
  ggplot(aes(year, lifeExp)) +
  geom_point() +
  geom_smooth(method = "lm", 
              formula = y ~ poly(x, 2), 
              se = FALSE)


```
