---
title: 'STAT 545 L3: Authoring'
author: "Jerry"
date: "2022-09-15"
output: 
  html_document:
    theme: cerulean
    toc: true 
    toc_float: true
---

# Authoring

## Intro to Markdown

Markdown is ideal for creating reproducible workflows. What this mean is that it is basically a workflow that makes re-running your analysis so simple that you can basically do it with one click.

We are going through basic formatting in a markdown file in the file `demo_markdown_file.md`

Also note, you can write comments that don't show up on the .html file with the following syntax:\
<!--- I'm a comment --->\
You should see an empty space above me.

### Writing PDF Files

Usually, you need a PDF writer such as LaTeX, but you we don't want to download it so we can use the R package "tinytex"

```{r}
#install.packages("tinytex")
#tinytex::install_tinytex()
```

## R Markdown Files

For R markdown, you need to download the "rmarkdown" package. Let's also download the "DT" package since we will need it in the future.

```{r}
#install.packages("rmarkdown")
#install.packages("DT")
```

FYI, everything that works in a basic markdown file also works in Rmd!

So what's the difference?\
- For one, you can write in code chunks in Rmd!\
- Rmds also come default with a YAML header, denoted between three dashes "---"\
- YAML headers are more or less metadata for Rmd and also gives info on how to format the doc.  
- Also, instead of previewing, Rmd uses the knitr package to knit these Rmds into HTML docs or others!  

When knitting Rmd files, it actually automatically runs your code and places the output tidily below. See for yourself by knitting this file!
```{r}
summary(cars)
```













